Crime in Los Angeles: Homicide Increases in the COVID-19 Era

### 2025-04-16 Slides/SlidesFinal.qmd
final slides version

---
title: "Crime in Los Angeles: Homicide Increases in the COVID-19 Era"
subtitle: "Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine"
format:
  revealjs:
# embed-resources: true
    slide-number: true
    chalkboard:
      buttons: false
    preview-links: auto
    css: styles.css
    footer: Group 5 - STAT 405/605 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Department of Statistics, Rice University - Source [quarto.org/revealjs](https://quarto.org/docs/presentations/revealjs/)
server: shiny
---

@@

```{r}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

```{r, results='asis'}
library(plotly)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

```{r, echo=FALSE}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

## Geographical Trends in Crime Rates By LAPD Bureau
```{r}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

## Correlation Heat Map
```{r}
library(lubridate)
# creating dataset for correlations
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions.
# covid data
query <- dbSendQuery(con, "
  SELECT 
    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
   cases_14day
  FROM covid_cases
")
covid_14day <- dbFetch(query,n=-1)
dbClearResult(query)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_14day$lagged <- covid_14day$date + 7

@@

@@

## Homicides Have Increased in the Pandemic

\begin{equation}
\text{HomicideCount}_i = \beta_0 + \beta_1 \text{year}_i + \epsilon_i
\end{equation}

Year is significant at a 0.01 level, but it has caused a very small increase, almost unnoticeable on the graph. Each year, homicide rates increase by 0.05 on average.

>>>>

---
title: "Crime in Los Angeles: Homicide Increases in the COVID-19 Era"
subtitle: "Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine"
format:
  revealjs:
# embed-resources: true
    slide-number: true
    chalkboard:
      buttons: false
    preview-links: auto
    css: styles.css
    footer: Group 5 - STAT 405 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Crime in Los Angeles
server: shiny
---

@@

```{r}
#| fig.align: 'center'
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

```{r, results='asis'}
#| fig.align: 'center'
library(plotly)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

```{r, echo=FALSE}
#| fig.align: 'center'
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

## Geographical Trends in Crime Rates By LAPD Bureau
```{r}
#| fig.align: 'center'
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

@@

## Correlation Heat Map
```{r}
#| fig.align: 'center'
library(lubridate)
# creating dataset for correlations
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions.
# covid data
query <- dbSendQuery(con, "
  SELECT 
    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
   cases_14day
  FROM covid_cases
")
covid_14day <- dbFetch(query,n=-1)
dbClearResult(query)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_14day$lagged <- covid_14day$date + 7

@@

Homicide rates are higher in the southeastern parts of Los Angeles.

@@

## Homicides Increased In Pandemic Years

\small
\begin{equation}
\text{HomicideCount}_i = \beta_0 + \beta_1 \text{year}_i + \epsilon_i
\end{equation}

```{r}
#| fig.align: 'center'
#| fig.width: 6
#| fig.height: 4
avg <- homicides_lat_long_yr %>% group_by(year) %>% summarize(avgHC = mean(HomicideCount)) %>% ungroup()
ggplot(avg, aes(x = year, y = avgHC)) +
  geom_line(color = "#7e03a8") +
  ggthemes::theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        plot.caption = element_text(size = 8)) +
  labs(y = "Average Homicide Count",
       x = "Year",
       title = "Average Homicide Count By Year",
       caption = "Average was derived from latitude and longitude groups used in the model.")
```
